---
title: "MEA_transmission"
output: html_document
date: "2024-08-02"
editor_options: 
  chunk_output_type: console
---

# 0. Libraries and functions
```{r}
home = getwd()
#
library("rhdf5")
library("tidyr")
library("dplyr")
library("ggplot2")
library("gsignal")
library("magick")
library("reticulate")
library("ggpubr")
# library instructions browseVignettes("rhdf5")

pick.stream <- function(trace0, trace1){
  # Pick the stream with the data
  traces.all = NULL
  answer <- readline (prompt="Which stream has the data? Enter [0] or [1] ")
  while ( nchar(answer) > 0){ #answer != "0" && answer != "1"
    
    if (answer == "0") {
      # put in dataframe
      traces.all = data.frame("time" = time0, traces0)
      print("Stream 0 saved")
      # trace.plot = gather(traces.all, "e", "V", 2:ncol(traces.all))
      # plot <- ggplot(trace.plot, aes(time, V, col = e) ) +
      #         geom_line()
      # print(plot)
      break
    } else if (answer == "1") {
      traces.all = data.frame("time" = time1, traces1)
      print("Stream 1 saved")
      #trace.plot = gather(traces.all, "e", "V", 2:ncol(traces.all))
      #plot <- ggplot(trace.plot, aes(time, V, col = e) ) +
              geom_line()
      #print(plot)
      break
    } else {
      print("Not valid input")
      answer <- readline (prompt="Enter [0] or [1] to pick a stream ")
    }
  } 
  return(traces.all)
} # still need this?
normalize.trace <- function( trace.in, bl.endFrame ){
  baseline = mean( trace.in[1:bl.endFrame] )
  trace.out = trace.in - baseline
  return(trace.out)
}
save.ask <- function() {
  answer <- readline (prompt="Press [y] to save; [n] to skip    ")
  while ( nchar(answer) > 0 ){
    if (answer == "y") {
      return("y")
      break
    } else if (answer == "n") {
      return("n")
      break
    } else {
      print("Not valid input")
      answer <- readline (prompt="Press [y] to save; [n] to skip ")
    }
  } 
}
calculateslope <- function(vector) {
  slopetrace = NULL
  for( i in 2:length(vector)){
    slopei = vector[i]-vector[i-1]
    slopetrace = append(slopetrace, slopei)  
  }
  slopetrace = append(slopetrace, slopetrace[length(slopetrace)])
  return(slopetrace)
}
encode_MEA <- function(){
    # Determine the arbitrary order of electrodes in the 32-MEA. 
  MEA_E6basipetal = list("rowX" = c("E1","E2","E3","E4","E5","E17","E28","E29","E30","E31","E32"),
                         "rowY" = c("E6","E7","E8","E9","E10",NaN,"E23","E24","E25","E26","E27"),
                         "rowZ" = c("E11","E12","E13","E14","E15","E16","E18","E19","E20","E21","E22") )
  # is it is rotated 180 degrees, the organization is inverted in x and y axes. 
  MEA_E6acropetal = rev(MEA_E6basipetal)
  MEA_E6acropetal[[1]] = rev(MEA_E6acropetal[[1]])
  MEA_E6acropetal[[2]] = rev(MEA_E6acropetal[[2]])
  MEA_E6acropetal[[3]] = rev(MEA_E6acropetal[[3]])
  # As convention, the A1,B1,C1 will correspond to the electrodes more basipetal. NEXT pick which disposition of the MEA was used
  # ask whether E6 is basipetal or acropetal
  cat("Is E6 acropetal (a) or basipetal (b)?")
  answer <- readline()
        while ( nchar(answer) > 0 ){
          if (answer == "a") {
            MEA = MEA_E6acropetal
            break
          }
          else if (answer == "b") {
            MEA = MEA_E6basipetal
            break
          }
          else {
            print("Not valid input. ")
            answer <- readline (prompt="Is E6 acropetal (a) or basipetal (b)?")
          }
        } 

  # Initialize an empty list to store key-value pairs
  arraykey1 <- list()
  # Generate key-value pairs
  rows = c("A","B","C")
  row_count = 1
  for (r in names(MEA) ){
    for (i in 1:11) {
      key <- MEA[[r]][i]  # Generate the corresponding value
      value <- paste0(rows[row_count],"_", i)  # Concatenate "A" with the number
      arraykey1[[key]] <- value  # Add the key-value pair to the list
    }
    row_count = row_count+1
  }
  return(arraykey1)
}
omit_electrodes <- function(array){
  cat("Remove electrodes? type X_#,Y_#... or (n) to omit ")
  answer <- readline()
        while ( nchar(answer) > 0 ){
          if (answer == "n") {
            break
          }
          else {
            remove.electrodes = as.vector( strsplit(answer,",") )[[1]]
            print("removed electrodes " )
            print(remove.electrodes )
            remove.rows = which( ( unique(array$electrode) %in% remove.electrodes)==TRUE )
            array[remove.rows,2:ncol(array)] = NaN
            break
          }
        }
  return(array)
}
```

The name of the .h5 raw files should have the following structure, which is mostly the automatic output of the MCS software.
yyyy-mm-ddThh-mm-ssxxxx_D-00042.h5 
where: 
xxxx should be changed to the genotype of the plant recorded. Any extra information in the file name should be erase.

1. Open and plot .h5 raw data 

# 1.1 Setup data manipulation specifications
input - Manually write here specifications of the analysis
output- creates a folder where to save all the analyzed files and plots
```{r}
foldername = "Output"
batchfolder = paste(home,foldername,sep = "/")
#downsample.to = 100 # provide the Hz you want your data to be downsampled to. NULL for skip downsample.
stim.time = 90 # seconds at which wound is done.
if ( !dir.exists(batchfolder) ){
  dir.create(batchfolder)
}
samplerate = 100
```


# 1.2 Run the MCS_dataManagement.py. Opens, downsamples to 100 Hz and saves.
/Users/j/Documents/GitHub/MCS_dataAnalysis/MCS_dataManager.py
input   - All the ".h5" files in the home folder of this project.
output  - a "Python-analysis" folder with ">savename<.csv" files with units s and mV and sample rate = 100 Hz.

run in the console: python /Users/j/Documents/GitHub/MCS_dataAnalysis/MCS_dataManager.py

# 1.3 Normalize to baseline, set array orientation, plot and save
input   - The "Python-analysis" with ">savename<.csv" files from 1.2
output  - Change the name of the "Python-analysis" folder to >batchfolder<.There, plot overlay of all the electrodes as ">savename<.pdf"
```{r}
setwd(home)
if ( dir.exists("Python-analysis") ){
  file.rename("Python-analysis", foldername)
} else {
  print("No Python-analysis folder")
}
setwd(batchfolder)
files = list.files()[grep(".csv", list.files())]
for (f in files){
  print(f)
  traces.all = read.csv(f)
  savename = substr(f,1,nchar(f)-4)
  genotype = "Col0" #substr(f,20,nchar(f)-4)
  
          # change name of electrodes by convention code
  # determine orientation of the array on the leaf
  MEA_key = encode_MEA()
  names(traces.all) = chartr(old = "X",new = "E",x = names(traces.all) ) # standardize nomenclature
  
  if ( is.null(MEA_key[[ names(traces.all)[1] ]] ) == FALSE ){
    for (i in 2:33 ) {
      names(traces.all)[i] = MEA_key[[ names(traces.all)[i] ]] 
    }
  }

  # go through all the columns and normalize the voltage to their baseline. That is the first 5 seconds of the recording
  normalize.point = which(traces.all$time == 5) # index at second 5
           
  # Now normalize
  for( c in 2:33){
    traces.all[,c] = traces.all[,c]-traces.all[normalize.point,c]
  }
  # save genotype and other data in the dataframe
  traces.all$genotype = genotype
  write.csv(traces.all, paste(savename,".csv",sep = ""), row.names = FALSE)
    # downsample for plotting.
  trace.plot = NULL
  for( c in 1:33 ){
    trace.plot =  cbind( trace.plot, downsample( traces.all[,c] , 10 ) )
  }
  trace.plot = data.frame(trace.plot)
  names(trace.plot) = names(traces.all)[1:33]
  trace.plot = gather( na.omit(trace.plot) , "e", "mV", 2:33)
  trace.events = traces.all$time[which(traces.all$events == 1)] - 0.5
  
# Plot
  plot <- ggplot(trace.plot, aes(time, mV, col = e) ) +
    geom_line() +
    geom_vline(xintercept = trace.events ) + 
    ggtitle(savename)
  print(plot)
  #if( save.ask() == "y"){
    ggsave(paste(savename,".pdf",sep = ""), plot = plot)
  #}
}
setwd(home) 
```



# 2 Plotting of transmission experiments with sine wave or ACR1 stimulation
```{r}
setwd(home)
#
power = "sineWave"
leaf.state = "Severed"
stimLeaf = "66"
distalLeaf = "64"

simDelay = 0.5

# Plot stimulate leaf
savename = list.files()[grep(stimLeaf, list.files() )] 
stimLeaf.all = read.csv(savename )
stimLeaf.plot = NULL
for( c in 1:33 ){
  stimLeaf.plot =  cbind( stimLeaf.plot, downsample( stimLeaf.all[,c] , 10 ) )
}
stimLeaf.plot = data.frame(stimLeaf.plot)
names(stimLeaf.plot) = names(traces.all)[1:33]
stimLeaf.plot = gather( na.omit(stimLeaf.plot) , "e", "mV", 2:33)
stimLeaf.events = stimLeaf.all$time[which(stimLeaf.all$events == 1)] - simDelay
if(power == "power-6") {
  # ymin = min(stimLeaf.plot$mV) - 1
  # ymax = max(stimLeaf.plot$mV) + 1
}
splot <- stimLeaf.plot%>% ggplot( aes(time, mV, col = e) ) +
  geom_line() +
  geom_vline(xintercept = stimLeaf.events ) + 
  ylim(ymin, ymax) +
  theme(legend.position = "none") +
  ggtitle(paste("Stimulated leaf - ",power,"_", leaf.state,sep = ""))

# Plot distal leaf
distalLeaf.all = read.csv( list.files()[grep(distalLeaf, list.files() )] )
distalLeaf.plot = NULL
for( c in 1:33 ){
  distalLeaf.plot =  cbind( distalLeaf.plot, downsample( distalLeaf.all[,c] , 10 ) )
}
distalLeaf.plot = data.frame(distalLeaf.plot)
names(distalLeaf.plot) = names(traces.all)[1:33]
distalLeaf.plot = gather( na.omit(distalLeaf.plot) , "e", "mV", 2:33)
distalLeaf.events = distalLeaf.all$time[which(distalLeaf.all$events == 1)] - simDelay

dplot <- distalLeaf.plot%>% ggplot( aes(time, mV, col = e) ) +
  geom_line() +
  geom_vline(xintercept = distalLeaf.events ) + 
  ylim(ymin, ymax) + 
  theme(legend.position = "none") +
  ggtitle(paste("Distal leaf - ",power,"_", leaf.state, sep = ""))

# mix both plots
savename = substr(savename,1, nchar(d.savename)-8)
ggarrange(splot, dplot, ncol = 1, nrow = 2)
ggsave(paste(power,leaf.state,savename,"pdf",sep = "."), width = 6, height = 5)

```

